### 2023-10-27 Joko-testing-implementasi-sistem-ujian/Ujian_ilman-samhabib_91122010.tex
ujian embedded system

\documentclass[12pt]{article}

\usepackage[a4paper,left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage[bahasa]{babel}

\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tikz}
% \usepackage{biblatex}
% \addbibresource{main.bib}
\usepackage{apacite}
\usepackage{longtable}
% \usepackage[backend=biber,style=apa,citestyle=apa,sorting=ynt]{biblatex}
% \addbibresource{main.bib}
\usepackage{usebib}
\usepackage{enumitem}

\bibinput{main}
\graphicspath{ {./images/} }
% \title{Tugas SPK}
% \author{Ilman Samhabib 91122010\\62/MMSI/SIB}
% \date{\today}

\begin{document}

% \maketitle
% \titlepage
% \newpage
% \tableofcontents
% \newpage
\begin{center}
   \textbf{Ujian: Testing dan Implementasi Sistem}
%    \\ Tema: \textbf{Ujian}
   \\Ilman Samhabib 9911220\textbf{10} 62/MMSI/SIB

\end{center}

\section*{Soal 1}

\begin{center}
   \includegraphics[width=17cm, height=9cm]{images/part-1.jpg}
\end{center}
\begin{center}
   \includegraphics[width=17cm, height=9cm]{images/part-2.jpg}
\end{center}
\begin{center}
   \includegraphics[width=17cm, height=9cm]{images/part-3.jpg}
\end{center}

\section*{Soal 2}

\begin{center}
   \includegraphics[width=17cm, height=9cm]{images/part-4.jpg}
\end{center}
\begin{center}
   \includegraphics[width=17cm, height=9cm]{images/part-5.jpg}
\end{center}
\end{document}